\documentclass{report}

\usepackage[vmargin=0.5in, hmargin=0.8in]{geometry}

%space between lines
\usepackage{setspace}
\setstretch{1.15}

\usepackage{minted} % for SQL code
\usepackage{parskip} % skip paragraph identation

\begin{document}

\chapter{Introduction to Relational Databases in SQL}
\section{Uniquely identify records with key constraints}
\subsection{Keys and superkeys}

What is a key?
\begin{itemize}
    \item Attributes(s) that identify a record Uniquely.
    \item As long as attributes can be removed: \textbf{superkeys}.
    \item If no more attributes can be removed: minimal superkey or \textbf{key}.
\end{itemize}

\vspace{0.5em}
\begin{center}
\begin{tabular}{lllll}
\textbf{license\_no} & \textbf{serial\_no} & \textbf{make} & \textbf{model} & \textbf{year} \\
Texas ABC-739       & A69352              & Ford          & Mustang        & 2             \\
Florida TVP-347     & B43696              & Oldsmobile    & Cutlass        & 5             \\
New York MP0-22     & X83554              & Oldsmobile    & Delta          & 1             \\
California 432-TFY  & C43742              & Mercedes      & 190-D          & 99            \\
California RSK-629  & Y82935              & Toyota        & Camry          & 4             \\
Texas RSK-629       & U028365             & Jaguar        & XJS            & 4             \\
\end{tabular}
\end{center}

\emph{Note: Table adapted from Elmasri, Navathe (2011): Fundamentals of Database Systems, 6th Ed., Pearson}

SK1 = \{license\_no, serial\_no, make, model, year\} \\
SK2 = \{license\_no, serial\_no, make, model\} \\
SK3 = \{make, model, year\}, 
SK4 = \{license\_no, serial\_no\}, 
SKi, ..., SKn

A lot of possible superkeys with different combinations. But only four possible \textbf{keys}

K1 = \{license\_no\}; 
K2 = \{serial\_no\}; 
K3 = \{model\}; 
K4 = \{make, year\}

Here K1 to 3 only consist of one attribute. Removing either "make" or "year" from K4 would result in duplicates. Only one candidate key can be the chosen key.

A simple way of finding out whether a certain column(or combination) contains only unique values, wrap everything within the COUNT() function.

\inputminted[
    firstline=1,
    lastline=7,
    highlightlines={4-5},
]{SQL}{keys_and_superkeys.sql}

\section{Primary keys}
\begin{itemize}
    \item One primary key per database table, chosen from candidate keys
    \item Uniquely identifies records, e.g. for referencing in other tables
    \item Unique and not-null constraint both apply
    \item Primary keys are time-invariant: choose columns wisely
\end{itemize}

\subsection{Specifying primary keys}

\inputminted[
    firstline=10,
    lastline=28,
    highlightlines={11,17,27}
]{SQL}{keys_and_superkeys.sql}

\emph{Note: Taken from PostgreSQL documentation.}

\subsection{Adding primary key to existing table}

\inputminted[
    firstline=30,
    lastline=31
]{SQL}{keys_and_superkeys.sql}

\section{Surrogate keys}


\end{document}